\documentclass[12pt]{article}

\usepackage[explicit]{titlesec}

\titleformat{\section}{\normalfont\Large\bfseries}{}{0em}{Article \thesection. #1}

% \usepackage[active,tightpage]{preview}
% \renewcommand{\PreviewBorder}{1in}

\title{The Northern Steppes Bylaws}
% \date{December 27, 2022}

\usepackage{enumerate}
\usepackage{enumitem}
\usepackage[margin=1in]{geometry}
\usepackage{xcolor}
\usepackage{color,soul}
% \usepackage[margin=0.75in, right=3.25in, marginparwidth=2.5in, marginparsep=0.25in]{geometry}
\usepackage[markup=bfit, addedmarkup=it, deletedmarkup=sout, commentmarkup=footnote, authormarkup=none, defaultcolor=red]{changes}
% \usepackage[markup=bfit, addedmarkup=it, deletedmarkup=sout, commentmarkup=margin, authormarkup=none, todonotes={textwidth=2.5in}]{changes}
% \usepackage{marginnote}

\usepackage[skip=1pt plus1pt, indent=40pt]{parskip}
% \setcommentmarkup{\marginpar{#1}}

\newlist{level}{enumerate}{10}
\setlist[level]{label*=\arabic*.}
\setlist[level,1]{label = \arabic*.,ref = \emph{Article \thesection.\ Section\ \arabic*.}}

% \newcommand{\headItem}[1]{\global\isHeadtrue\item {\bfseries #1}}
% \newcommand{\maybeHeadItem}{%
%     \ifisHead\bfseries\global\isHeadfalse\fi%
% }
% \newif\ifisHead

\NewDocumentCommand{\subject}{ m }{%
    \begin{center}
        \bfseries%
        Re: #1%
    \end{center}
}%
\NewDocumentCommand{\name}{ m }{{\bfseries #1}}
\NewDocumentCommand{\email}{ m }{ #1 }

% \renewcommand{\thesection}{Article \arabic{section}}
% \renewcommand{\thesection}{Article \Roman{section}}

% \newcommand{\Newpage}{\end{preview}\begin{preview}}
% \newcommand{\newpart}[2][]{\colorbox{green!40}{#2}\expandafter\ifx\expandafter\relax\detokenize{#1}\relax\else\textnormal{\footnote{#1}}\fi}
\newcommand{\newpart}[2][]{\hl{#2}\expandafter\ifx\expandafter\relax\detokenize{#1}\relax\else\textnormal{\footnote{#1}}\fi}

\begin{document}
% \begin{preview}

% \maketitle

% \name{}

% \listofchanges

\section{Name and Purpose}
\begin{level}
    \item This organization shall be known as The Northern Steppes, a realm of Belegarth Medieval Combat Society.
    \item The purpose of this organization is to participate in the Foam Sport Belegarth, which includes running regularly scheduled practices and events.
\end{level}

\section{\newpart{Definitions}}
\begin{level}
    \item \textbf{Realm}. This organization, previously known as a \emph{chapter}.
    \item \textbf{Leadership}. Individuals that fulfill offices and roles defined in \emph{Article~\ref{leadership}}.
    \item \textbf{The Council}. The group of officers defined in \ref{the_council}
    \item \textbf{Officer}. An individual belonging to the \emph{Council} that fulfills an office defined in \ref{council_officers}
    \item \textbf{Role-Holder}. An individual that fulfills a Supplemental Role defined in \ref{supplemental_roles}
    \item \textbf{Members}. \emph{Members of the Realm}. Individuals that have registered via \emph{Article~\ref{membership}}.
    \item \textbf{Participants}. Anyone attending Realm-hosted events, both combatants \& non-combatants.
    \item \textbf{Event}. Any scheduled happening arranged by the Realm: weekly practices, build days, special events, member meetings.
    \item \textbf{Special Event}. A larger, fighting-based event, where fighters outside of the Realm are explicitly invited, e.g. GotC, Kill-Grill-Chills
    \item \textbf{Troll}. The check-in set up at special events where participants can pay, sign-in, and get information about the event.
    \item \textbf{Quorum}. The minimum number of people needed for a vote to be valid.
    \item \textbf{Simple Majority}. More \emph{Yes} votes than \emph{No} votes with the total number of \emph{Yes}/\emph{No} votes exceeding quorum.
    \item \textbf{Consensus Approval}. A quorum of \emph{Yes} votes with 0 \emph{No} votes.
    \item \textbf{Online}. Regarding voting, a group-specific Discord channel, group text, or group chat.
\end{level}

\section{Rule sets}
\begin{level}
    \item All realm-hosted events shall follow the Belegarth rule set established under the Book of War (BoW), as approved by the Belegarth War Council, with the addition of:
    \begin{level}
        \item \textbf{Age Limits}. In addition to the BoW rule of age limits, Children ages 12--15 that are approved by the Council after discussion with their parent(s)/guardian(s) may attend on the field practices only with their parent(s)/guardian(s) present after signing a physical waiver.
        \item \textbf{Electronic Wavier}. A fillable electronic version of the Northern Steppes' Belegarth waivers are made available online. These waivers are available only to those ages 18 and up.
        \item \textbf{Rigid Armor}. Rigid metal knee and elbow armor is allowed.
        \item \textbf{Healing Poems}. A poem of Healing must be recited or read aloud. The minimum length of a poem of healing is 180 syllables and must be approved at event check-in.
        \begin{level}
            \item The healer cannot heal \newpart[Inclusive language]{themself} \deleted{her/himself} or their own equipment.
            \item To heal a person, the Healer must be in direct physical contact with the injured person. The healer may only heal one person at a time.
            \item A healer may not heal a person while either are under direct attack. If the healer or injured person is attacked or distracted, or contact is broken for any reason, the entire healing poem must be read or recited again.
            \item When the healer completes reading the healing poem, all wounds and any items held by the wounded fighter are healed or repaired (i.e. armor, bow, shield, etc)
        \end{level}
    \end{level}
    \item \newpart{All leadership, members, and participants must abide by the Code of Conduct established by the Council.}
\end{level}

\section{Membership}\label{membership}
\begin{level}
    \item \textbf{Benefits}. While anyone may attend events and use loaner gear, participants may annually register as a \emph{Member} of the Northern Steppes and be granted the following benefits:
    \begin{level}
        \item \newpart{Priority use of loaner gear.}
        \item \newpart{Use of any Council-specified, Member-only loaner gear.}
        \item Voting rights during the annual meeting and other voting events.
        \item \newpart{Entrance to all Northern Steppes non-overnight events, such as:}
        \begin{level}
            \item A singular day at GotC (longer stays will require the difference to be paid).
            \item Build days that may have an entry fee.
        \end{level}
    \end{level}
    \item \textbf{Registration}. \newpart{Members may not register in the period between the closing of bylaw amendments mentioned in Article~\expandafter{\ref{amendments}}
        and the end-of-voting during the annual meeting}, and any dues paid during this period will cause their registration to be delayed until the end-of-voting during 
        the annual meeting. Otherwise, registration happens as soon as dues are paid.
    \item \textbf{Term}. Membership lasts from registration until the next end-of-voting at the annual meeting. 
    \item \textbf{Dues}. The standard annual rate is \$30.
    \begin{level}
        \item \newpart[Discount and/or standard rate may be adjusted---we could remove it from the bylaws \& let the Council decide it on a yearly basis.]{To encourage registering at the beginning of a season, any registration prior to May 1st will receive a discount of \$5.} 
        \item If dues are unaffordable by someone wanting to become a Member, the Council may approve an explicit amount of volunteering, completed prior to their registration, in lieu of dues.
    \end{level}
    \item \textbf{Uses}. Usage of funds is decided by the Council with information supplied by the Treasurer, and may include, but is not limited to:
    \begin{level}
        \item Event or site reservation costs and fees.
        \item Props or equipment needed for events.
        \item Website hosting/maintenance.
        \item Loaner gear creation, repair, and maintenance.
    \end{level}
    \item \textbf{Removal}. If a Member is banned from the realm, their membership is automatically revoked, with no refundment of dues.
\end{level}

\section{\newpart{Leadership}}\label{leadership}
\begin{level}
    \item Leadership consists of all Council officers and anyone fulfilling a supplemental role, and all are required to:
    \begin{level}
        \item Be members of the realm at the time of their election.
        \item Remain unpaid volunteers.
        \item Be clearly identifiable at realm-hosted special events.
    \end{level}
    \item \textbf{The Council} is the main decision-making entity of the realm.\label{the_council}
    \begin{level}
        \item \newpart{\textbf{Voting}. The Council may vote on day-to-day matters for which there is no clear consensus.}
        \begin{level}
            \item In Council votes, a quorum is three officers, and must include the President or Vice President. 
            \item If a quorum is present in person, an attempt should be made to discuss and gather the votes of those not present online, but the voting may happen in person.
            \begin{level}
                \item Otherwise, the discussion and voting should happen online.
            \end{level}
            \item All vote results should be announced to Leadership and recorded by the Secretary.
            \item The Council may choose to discuss any issue with Leadership-at-large or Realm Members, but only Council officers may vote.
        \end{level}
        \item \textbf{Responsibilities}:
        \begin{level}
            \item Welcome newcomers to the sport/realm.
            \item Consistently attend realm's events---practices, special events, build days, and meetings---for a recognizable portion of the duration. 
            \begin{level}
                \item Attending practice a minimum of every other week is imperative to maintain these positions. 
                \item Prolonged absences are to be allowed only for vacations, illness, and other unordinary circumstances. 
                \item If an officer does not show at scheduled practice for three consecutive weeks without communicating a written explanation to the Council or the realm as a whole, they are to be dismissed and a replacement is to be selected by the remaining officers. 
                \item Council officers who work are expected to adjust their schedules such that they can regularly attend practice and fulfill their commitment to the realm.
            \end{level}   
            \item Have regular access to the internet and email so they are aware of incoming email and group activity on forums (Facebook, Discord, et al).
            \item Have moderating privileges on the forums and are to take responsibility managing as needed.
            \item Meet for at least 15 minutes every month.
            \item Organize an annual meeting as described in Article~\ref{annual_meeting}.
        \end{level}
        \item \newpart{\textbf{Capabilities}}. A \emph{Simple Majority} is required unless otherwise specified:
        \begin{level}
            \item \textbf{Create a new supplemental role}, as long as (a) it has a title and explicit written responsibilities, and (b) realm members are notified.
            \item \textbf{Assign any role} (vacant or occupied) for the duration of a special event, as long as (a) that person is a realm member, and (b) special event attendees are notified, either ahead of time or during the special event.\label{temp_role}
            \item \textbf{Appoint any vacant role} with \emph{Consensus Approval}, as long as (a) that person is a realm member, and (b) realm members are notified.
            \item \textbf{Vacate any role} (with responsibilities returning to the officer described) with \emph{Consensus Approval}, and notify realm members.
            \item \textbf{Appoint any vacant office} with \emph{Consensus Approval}, as long as (a) that person is a realm member, (b) that person isn't already an officer, and (c) realm members are notified.\label{officer_replacement}
            \item \textbf{Remove an officer} with \emph{Consensus Approval}, notify realm members, and replace the officer via Section~\ref{officer_replacement}\label{officer_removal}
            \item \textbf{Ban a participant} from the realm with a \emph{Consensus Approval}.
        \end{level}
        \item \textbf{Removal}. If a petition signed by \newpart{a quorum of realm members} \deleted{Abstained signatures are only considered ``nay'' in this scenario; this is in exception to Article VI Section 1.1.}\footnote{Revised quorum definition accomplishes the same goal} to impeach an officer is presented to the Council, the Council should:
        \begin{level}
            \item Give that officer one month to turn over responsibilities to another officer, at which point that office is considered vacant, and may be appointed via Section~\ref{officer_replacement}\label{officer_yield}
            \item \newpart{\emph{OR} vote to remove the officer immediately via Section~\expandafter{\ref{officer_removal}}. If the vote fails, Section~\expandafter{\ref{officer_yield}} must be followed.}
        \end{level}
        \item \textbf{Resignation}. An officer may resign at any time, and should be replaced via Section~\ref{officer_replacement}
    \end{level}
    \item \textbf{Council Officers}:\label{council_officers}
    \begin{level}
    \item \textbf{President}
        \begin{level}
            \item Serve as coordinator of the realm and the Council.
            \item Plan for the future of the realm. 
            \item Outline what needs done for recruitment, event planning, etc. then assign people to do it individually or through committees/teams.
            \item Work towards obtaining tax exempt status.
            \item Work with the Treasurer to file all tax returns and reports required to maintain tax exempt status.
            \item Establish standard operating procedures to:
            \begin{level}
                \item Run the annual meeting
                \item Handle the annual election or any other voting matter
            \end{level}
        \end{level}
        \item \textbf{Vice President}
        \begin{level}
            \item Fulfill the duties of the President in their absence, or as delegated.
            \item Fill the gaps in duties described in these bylaws.
            \item Encouraged to also run for a supplemental role.
        \end{level}
        \item \textbf{Secretary}
        \begin{level}
            \item Waiver gathering and tracking.
            \item Update realm information as needed.
            \item Keep record of former information.
            \item Operate Troll at special events, or train someone to do so.
            \begin{level}
                \item \newpart{Keep a list of people that are able to operate Troll and make Leadership aware of them.}
            \end{level}
            \item Take notes/minutes during meetings.
            \item \newpart{Store/update any standard operating procedures.}
            \item Establish standard operating procedures to:
            \begin{level}
                \item Operate Troll at special events.
            \end{level}
        \end{level}
        \item \textbf{Event Officer}
        \begin{level}
            \item Plan and schedule realm events, delegating duties where necessary:
            \begin{level}
                \item Weekly practices
                \item Special events, such as:
                \begin{level}
                    \item Longer day or overnight/multi-day events
                    \item “Kill-Grill-Chill” potluck-type events
                \end{level}
                \item Build days
                \item Annual meetings
            \end{level}
            \item In collaboration with the Public Relations Coordinator, inform realm members and potential attendees about realm events.
            \item The Council is to be consulted on any subgroup sponsorship when budgeting and planning for the realm event.
            \item \newpart{Coordinate a presence at other events, and serve as/delegate a campmaster.}
            \item \newpart{Establish standard operating procedures for event-specific tasks, such as:}
            \begin{level}
                \item Opening/setting-up.
                \item Daily happenings \& officer responsibilities during the event.
                \item Closing/cleaning-up.
            \end{level}
        \end{level}
        \item \textbf{Safety Officer}
        \begin{level}
            \item Be familiar with the Belegarth Book of War, and all realm-specific rules.
            \item Educate participants on rules and safety, including realm-specific rules. 
            \item Head security at special events.
            \item Oversee weapons-check at special events.
            \item Test and approve weapons at practices.
            \item \newpart{Establish standard operating procedures to:}
            \begin{level}
                \item Marshal battles, including the training of volunteers.
                \item Check weapons, including the training of volunteers.
            \end{level}
        \end{level}
\end{level}
\item \textbf{Supplemental Roles} are elected or appointed, and default to a specific Council officer that inherits the responsibilities if the role is vacant. A Council officer or realm member may have multiple roles.\label{supplemental_roles}
\begin{level}
    \item A realm member may be assigned in an acting capacity for the duration of a special event by the Council via Section~\ref{temp_role} if the role is vacant or the existing holder is otherwise unable to perform their duties.
    \item \textbf{Treasurer}. Vice President if vacant. Has the following responsibilities:
    \begin{level}
        \item Accurately record all financial records.
        \item Collect all receipts.
        \item File any appropriate tax information as needed.
        \item Prepare an annual financial report for public release.
        \item Establish and maintain a checking account.
        \item Review and remit payment for debts incurred.
        \item Establish standard operating procedures to:
        \begin{level}
            \item File taxes
            \item Reimburse purchases
        \end{level}
    \end{level}
    \item \newpart{\textbf{Quartermaster}}. Safety Officer if vacant. Has the following responsibilities:
    \begin{level}
        \item Be knowledgeable on gear construction techniques.
        \item Maintain loaner gear and replace/repair as needed.
        \item Store or designate an officer to store loaner gear, which should be brought to all fighting-based events.
        \item Attend and lead realm build-days, and in collaboration with the Event Officer, plan them.
        \item Establish instructions for recommended weapon construction.
    \end{level}
    \item \newpart{\textbf{Battlemaster}}. Event Officer if vacant. Has the following responsibilities:
    \begin{level}
        \item Plan, design, and construct field battles for fighting-based events, in collaboration with the Event Officer.
        \item Assign and coordinate marshals, after marshals are trained by the Safety Officer.
        \item Serve as or appoint a head marshal, who will:
        \begin{level}
            \item Run field battles and rally fighters to the field.
            \item Schedule breaks during battles to ensure fighters get adequate hydration, nutrition, and cooling.
        \end{level}
    \end{level}
    \item \textbf{Public Relations Coordinator}. President if vacant. Has the following responsibilities:
    \begin{level}
        \item Main point of contact for all realm-level media coverage.
        \item Have access to and regularly check the realm email account.
        \item Manage all social media channels officially associated with the realm.
        \item Create and post content on social media platforms.
        \item Oversee outreach campaigns (physical, social media, email, etc) such as:
        \begin{level}
            \item Dues renewal reminder
            \item Event announcement \& promotion
            \item General recruiting
            \item Election and voting motions
        \end{level}
        \item Work to increase social media followers for the realm.
        \item \newpart{Establish standard operating procedures to:}
        \begin{level}
            \item Create content.
            \item Launch social media/email campaigns.
        \end{level}
    \end{level}
    \item \textbf{Webmaster}. Secretary if vacant. Has the following responsibilities:
    \begin{level}
        \item Maintain the realm website, making any and all changes, as mandated by the Council.
        \item In collaboration with the Event Officer, update special event pages on the website.
        \item In collaboration with the Secretary, maintain an email list for realm members, special event attendees, etc.
        \item In collaboration with the Public Relations Coordinator, facilitate email campaigns.
        \item \newpart{In collaboration with the President, communicate voting efforts.}
    \end{level}
    \item \textbf{Belegarth War Council Representative (2)}. President (primary) and Vice President (alternate) if vacant. Has the following responsibilities:
    \begin{level}
        \item Represent the realm in the Belegarth War Council.
        \item Attend outside special events in addition to realm events.
        \item Communicate with the Council and act as national liaisons to the realm, facilitating the will of the realm in an unbiased manner.
        \item Must have the means to access the internet at least one per week.
        \item May not serve as a National Belegarth Officer.
        \item Strive to be respondent, communicating any abstains in voting to the Council.
    \end{level}
\end{level}

\end{level}

\section{Meetings and Quorum}\label{annual_meeting}
\begin{level} 
    \item \newpart{A realm member meeting may be scheduled at any time, given that members are notified at least four (4) weeks prior to the meeting.}
    \item A single annual meeting is required for realm members to elect new leadership. 
    \begin{level}
        \item The date, time, and location is determined by the Council. 
        \item This should be completed before the end of January each year.
    \end{level}
    \item During any member meeting, \newpart[]{the quorum shall consist of 50\% of members plus one (1)}. \deleted{Quorum is defined as 75\% of the subject in question. This applies to all votes and attendance of meetings. Regarding votes, a vote of “abstain” is not included in the total count of those present for that issue}\footnote{This wording was confusing, and as our membership grows, it may be impossible to get this many people. 50\% + 1 is safer, and will prevent minority rule.}
\end{level}

\section{Amendments}\label{amendments}
\begin{level} 
    \item Amendments to this document shall be introduced to the Council in writing, no less than three (3) weeks prior to a member meeting. 
    \item Amendments should be read and discussed during a member meeting prior to voting, if quorum is present.
    \item \newpart{A \emph{Simple Majority} is required for an amendment to pass.} \deleted{Quorum of those in attendance is necessary for them to pass.}\footnote{This wording was unclear, as quorum is usually applied to the number of votes needed for a valid vote, not pass/fail status.}
\end{level}

\section{Enabling Act}
\begin{level} 
    \item This set of Bylaws is officially ratified when approved through the current voting membership.
\end{level}

% \end{preview}
\end{document}